---
title: "Predicting Power Usage with a Random Forest Model"
format: html
editor: visual
---

```{r}
# Necessary libraries
library(dplyr)
library(lubridate)
library(tidyr)
library(ranger)
library(xgboost)
library(ggplot2)
df_train <- read.csv("./data/df_train.csv")
df_test <- read.csv("./data/df_test.csv")
```

## Preparation

Preview the data:

```{r}
glimpse(df_train)
glimpse(df_test)
```

Define `prepare_features` to format test and train data sets and one-hot encode categorical variables for the gradient-boosted tree:

```{r}
prepare_features <- function(df, fmla) {
  df <- df %>%
    select(-year) %>%
    mutate(date = as.Date(date, format = "%m/%d/%Y")) %>%
    mutate(across(c(semester, quarter, day_in_week, week_in_year, day_in_year, month), as.factor))
  encoded_df <- as.data.frame(model.matrix(fmla, data = df))
  df <- df %>%
    bind_cols(encoded_df)
  return(df)
}
```

Format and extract features from the test and train data:

```{r}
fmla <- as.formula("power_consumption ~ semester + quarter + day_in_week + week_in_year + day_in_year + month - 1")

df_test <- prepare_features(df_test, fmla)
df_train <- prepare_features(df_train, fmla)

df_train_features <- df_train %>%
  select(-c(date, power_consumption, semester, quarter, day_in_week, week_in_year, day_in_year, month))

df_test_features <- df_test %>%
  select(-c(date, power_consumption, semester, quarter, day_in_week, week_in_year, day_in_year, month))

outcome_train <- df_train["power_consumption"]
outcome_test <- df_test["power_consumption"]
```

## Train Models

Train baseline linear model:

```{r}
model.lm <- lm(fmla, data = df_train)
summary(model.lm)
```

Train random forest (RF) model using `ranger`. `mtry` is set to `floor(p/3)` per the standard regression heuristic. OOB error serves as a built-in cross-validation estimate — no separate validation loop is needed:

```{r}
set.seed(1221)
model.rf <- ranger(formula = fmla,
                   data = df_train,
                   num.trees = 500)

df_test$pred.lm <- predict(model.lm, newdata = df_test)
df_test$pred.rf <- predict(model.rf, data = df_test)$predictions
```

## Gradient-Boosted Tree

We now train a gradient-boosted tree. Because the test and train data do not share the same one-hot encoded columns after `prepare_features`, only common features are used. Column alignment is performed before cross-validation so CV results reflect the actual feature set used for training:

```{r}
common_cols <- intersect(colnames(df_test_features), colnames(df_train_features))
df_train_features <- subset(df_train_features, select = common_cols)
df_test_features <- subset(df_test_features, select = common_cols)
```

Run cross-validation to find the optimal number of boosting rounds:

```{r}
set.seed(126)

params <- xgb.params(
  objective = "reg:squarederror",
  eta = 0.3,
  max_depth = 5
)

dtrain <- xgb.DMatrix(
  data  = as.matrix(df_train_features),
  label = as.matrix(outcome_train)
)

cross_val <- xgb.cv(
  params = params,
  data = dtrain,
  nrounds = 100,
  nfold = 5
)

eval_log <- cross_val$evaluation_log
best_nrounds <- eval_log[which.min(eval_log$test_rmse_mean), "iter"][[1]]

eval_log %>%
  summarize(ntrees.train = which.min(train_rmse_mean),
            ntrees.test = which.min(test_rmse_mean))
```

Train the gradient-boosted tree using the optimal number of rounds from cross-validation:

```{r}
set.seed(123)
model.xgb <- xgboost(data = as.matrix(df_train_features),
                     label = as.matrix(outcome_train),
                     nrounds = best_nrounds,
                     objective = "reg:squarederror",
                     eta = 0.3,
                     max_depth = 5,
                     verbose = FALSE
)

df_test$pred.xgb <- predict(model.xgb, newdata = as.matrix(df_test_features))
```

## Model Comparison

Compare RMSE across all three models. The random forest OOB RMSE is its built-in cross-validation estimate:

```{r}
rmse_lm <- df_test %>%
  mutate(residual = pred.lm - power_consumption) %>%
  summarize(rmse = sqrt(mean(residual^2))) %>%
  pull(rmse)

rmse_rf <- df_test %>%
  mutate(residual = pred.rf - power_consumption) %>%
  summarize(rmse = sqrt(mean(residual^2))) %>%
  pull(rmse)

rmse_rf_oob <- sqrt(model.rf$prediction.error)

rmse_xgb <- df_test %>%
  mutate(residual = pred.xgb - power_consumption) %>%
  summarize(rmse = sqrt(mean(residual^2))) %>%
  pull(rmse)

tibble(
  model = c("Linear", "Random Forest (test)", "Random Forest (OOB)", "XGBoost"),
  rmse  = c(rmse_lm, rmse_rf, rmse_rf_oob, rmse_xgb)
) %>% arrange(rmse)
```

Plot the models (colored) against test data (black):

```{r}
plot <- df_test %>%
  pivot_longer(cols = c(pred.xgb, pred.lm, pred.rf), names_to = "model", values_to = "predicted") %>%
  ggplot() +
  geom_line(aes(x = date, y = power_consumption), alpha = 0.7) +
  geom_line(aes(x = date, y = predicted, color = model)) +
  labs(x = "Date", y = "Power Consumption (kwH)") +
  theme_minimal() +
  ggtitle("Power Consumption Over Time")
plot
```
